---
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
---

# RNAseq downstream analysis            

```{r message=FALSE, warning=FALSE, include=FALSE}
#necessary packages and functions  
source("I:/Shared_win/projects/RNA_normal/analysis.r")
```
         
          
## load expression mtx        

already filtered in 'all_initial' part             

```{r echo=FALSE}
dat = './RNAseq.SS2_LY_20210608_mod20220117.filt_cpm.tissues.pc_gene.csv'
mtxname = 'SS2_LY_20210608.tissues'

# specify names of 'A vs B', and cell idx of As and Bs
Aname = "SI" 
Bname = "nonSI"  

#Aidx = c()
#Bidx = c()

mat_pc <- read.table(dat, header = TRUE, stringsAsFactors = F, sep = ",")
rownames(mat_pc) <- mat_pc$gene
mat_pc <- mat_pc[,2:ncol(mat_pc)]
mat_pc <- as.matrix(mat_pc)

Aidx = grep("SI",colnames(mat_pc))
Bidx = grep("SI",colnames(mat_pc),invert = T)

# tpm
datt = './RNAseq.SS2_LY_20210608_mod20220117.filt_tpm.tissues.pc_gene.csv'

matt_pc <- read.table(datt, header = TRUE, stringsAsFactors = F, sep = ",")
rownames(matt_pc) <- matt_pc$gene
matt_pc <- matt_pc[,2:ncol(matt_pc)]
matt_pc <- as.matrix(matt_pc)
```

check filtered pc(protein coding) matrix        
```{r}
dim(mat_pc[,c(Aidx,Bidx)])
head(mat_pc[,c(Aidx,Bidx)])
``` 

## QC  

design info  
```{r echo=FALSE, paged.print=FALSE}
design <- data.frame(condition=factor(c(rep(Aname,length(Aidx)),rep(Bname,length(Bidx)))),
                     replicate=factor(colnames(mat_pc[,c(Aidx,Bidx)])))
rownames(design) <- colnames(mat_pc[,c(Aidx,Bidx)])
design$condition <- factor(design$condition, levels = c(Aname,Bname))
design
```

### condition mean expression             

```{r cross2, echo=FALSE, fig.height=6, fig.width=6, warning=FALSE}
plot(data.frame(Aname=rowMeans(mat_pc[,Aidx]),
                Bname=rowMeans(mat_pc[,Bidx])), pch=20, log = 'xy',
     xlab = paste0("Mean normalized counts (",Aname,")"),xlim=c(0.1,100000),
     ylab = paste0("Mean normalized counts (",Bname,")"),ylim=c(0.1,100000),
     main = paste0(dim(mat_pc)[1]," filtered protein coding genes "))
```

### PCA          

```{r include=FALSE}
rv <- rowVars(mat_pc[,c(Aidx,Bidx)])
selt <- order(rv, decreasing = TRUE)[seq_len(2000)]
pca2 <- stats::prcomp(t(mat_pc[selt,c(Aidx,Bidx)]), scale.=TRUE, center= TRUE)
#pca2$x

pca_d <- as.data.frame(pca2$x)

pca_d[,"condition"] = design$condition
pca_d[,"batch"] = rownames(design)
pca_d[,"replicate"] = design$replicate
```

```{r echo=FALSE}
plot(pca2$sdev,pch=20, xlab="PC",ylab="Standard Variation", main="PCs of top2000 highly variable genes")
```

##### PCA summary                 

```{r echo=FALSE}
# variance
summary(pca2)
```

```{r echo=FALSE, fig.height=7.2, fig.width=8, message=FALSE, warning=FALSE}
#plot the PCA
#create a classic 2-dimension PCA plot (first two principal components) with conditions and library methods indicated
cols <- c("#481567FF","#1F968BFF","#B5640B")

cols <- c(cols[3], cols[2])
names(cols) <- c(as.character(unique(design$condition)[1]),as.character(unique(design$condition)[2]))

centroids <- aggregate(cbind(PC1,PC2)~condition,pca_d,mean)
conf.rgn <- do.call(rbind,lapply(unique(pca_d$condition),function(t)
  data.frame(condition=as.character(t),
             ellipse::ellipse(cov(pca_d[pca_d$condition==t,1:2]),
                     centre=as.matrix(centroids[t,2:3]),
                     level=0.95),
             stringsAsFactors = FALSE)))


#p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition, shape=condition))
p1 = ggplot(data=pca_d, aes(x=PC1, y=PC2, color=condition))
p1 = p1 + geom_point(size=3.5)
p1 = p1 + ggrepel::geom_text_repel(mapping = aes(label=pca_d$batch),size=3.5, show.legend = F, max.overlaps = 20)
#p1 = p1 + stat_ellipse(type="norm", linetype=2)
p1 = p1 + geom_path(data=conf.rgn)

p1 = p1 + labs(title="PCA\nusing CPM(top2000 highly variable genes) ",
               color="Condition", shape="batch")
p1 = p1 + scale_colour_manual(values = cols) 
#p1 = p1 + scale_shape_manual(values=c(16,17,15))
p1
```

### tSNE       

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=8,fig.height=7.2}
set.seed(128)
PCs <- 1:10
PCss <- paste0(PCs[1],":",PCs[length(PCs)])
pca2.emb <- pca2$x %*% diag(pca2$sdev**2)
tsne_out <- Rtsne::Rtsne(
  as.matrix(pca2.emb[,PCs]),
  pca = FALSE,
  perplexity = 3,
  max_iter = 2000
)



#cols <- c("Meninges" = "#481567FF", "nonMeninges" = "#1F968BFF")
p6 = ggplot(data=data.frame(tsne_out$Y), aes(x=X1, y=X2)) + labs(x="tSNE_1",y="tSNE_2", seed=FALSE) + 
  #geom_point(aes(colour=design$condition, shape=design$condition),size=3.5) + 
  geom_point(aes(colour=design$condition),size=3.5) + 
  #stat_ellipse(aes(colour=design$condition),type="norm", linetype=2) + 
   ggrepel::geom_text_repel(mapping = aes(label=colnames(mat_pc[,c(Aidx,Bidx)])),size=3.5) + 
  scale_colour_manual(values = cols) + 
  #scale_shape_manual(values=c(16:10)) + 
   labs(title=paste0("tSNE\nusing PC",PCss),
               color="Condition", shape="batch")
p6
```

### correlation        

##### correlation (log2(CPM+1)) of all filtered protein-coding genes          
```{r echo=FALSE,message=FALSE,warning=FALSE,fig.width=5.2,fig.height=5.2}
  Ret_mat <- log2(mat_pc[,c(Aidx,Bidx)]+1)
  Ret_cor <- cor(Ret_mat)
  
  lower = 0.65
  upper = 1
  pal <- "Reds"
  
  ht1 <- ComplexHeatmap::Heatmap(Ret_cor, col = circlize::colorRamp2(seq(lower, upper, ((upper-lower)/7)),RColorBrewer::brewer.pal(8, pal)),
                  heatmap_legend_param = list(
                  color_bar = "continuous",
                  legend_direction = "horizontal",
                  legend_width = unit(5, "cm"),
                  title_position = "topcenter"),
                  name = "Pearson correlation",
                  column_names_gp = grid::gpar(fontsize = 10),
                  row_names_gp = grid::gpar(fontsize = 10),
                  top_annotation = NULL)
  heatplot1 <- ComplexHeatmap::draw(ht1, heatmap_legend_side = "top")
``` 

```{r eval=FALSE, include=FALSE}
pdf(file = "./figures.tissues/correlation.all_filtered_protein_coding.pdf",
    height = 5.2, width = 5.2)
heatplot1
dev.off()
```

##### correlation (log2(CPM+1)) of top2000 highly variable genes           

```{r echo=FALSE,message=FALSE,warning=FALSE,fig.width=5.2,fig.height=5.2}
  Ret_mat <- log2(mat_pc[selt,c(Aidx,Bidx)]+1)
  Ret_cor <- cor(Ret_mat)
  
  lower = 0.35
  upper = 1
  pal <- "Reds"
  
  ht1 <- ComplexHeatmap::Heatmap(Ret_cor, col = circlize::colorRamp2(seq(lower, upper, ((upper-lower)/7)),RColorBrewer::brewer.pal(8, pal)),
                  heatmap_legend_param = list(
                  color_bar = "continuous",
                  legend_direction = "horizontal",
                  legend_width = unit(5, "cm"),
                  title_position = "topcenter"),
                  name = "Pearson correlation",
                  column_names_gp = grid::gpar(fontsize = 10),
                  row_names_gp = grid::gpar(fontsize = 10),
                  top_annotation = NULL)
  heatplot2 <- ComplexHeatmap::draw(ht1, heatmap_legend_side = "top")
``` 

```{r eval=FALSE, include=FALSE}
pdf(file = "./figures.tissues/correlation.top2000_hvg.pdf",
    height = 5.2, width = 5.2)
heatplot2
dev.off()
```


## Differential Analysis                


```{r include=FALSE}
ret <- run_MEedgeR(MAT = mat_pc[,c(Aidx,Bidx)], ED = Aname, WT = Bname, n1 = length(Aidx),n2 = length(Bidx),
                         lcm = 0, padj = 0.05,lfc = log2(1.5), gsea = TRUE)
```


```{r}
ret$SUM
```

```{r paged.print=FALSE}
ret$gsea
```

```{r eval=FALSE, include=FALSE}
write.table(ret$gsea,
            paste0("figures.tissues/gsea_hallmark.DEGs_p0.05_FC1.5.",
                   Aname,"_vs_",Bname,".csv"),
            col.names = T, row.names = F, sep = ",", quote = F)
```


##### using cutoff: p.adjust <0.05, FC >2         
                      
```{r message=FALSE, warning=FALSE, include=FALSE}
rets <- finalplot(mat_pc[,c(Aidx,Bidx)],data.frame(ret$DEG_edgeR), paste0(Aname,"_vs_",Bname), 0.05, 1.5, 
                  Sign = FALSE, Sign_dn = 40, Sign_up = 40, padjust = TRUE)
```

```{r echo=FALSE}
cat(paste0(Aname," vs ",Bname," \nDEGs: "),length(c(rets$up,rets$down)),
    "\n  up: ",length(c(rets$up)),
    "\ndown: ",length(c(rets$down)))
```

##### correlation (log2(CPM+1)) of edgeR DEGs             
```{r echo=FALSE,message=FALSE,warning=FALSE,fig.width=5.2,fig.height=5.2}
    Ret_mat <- log2(mat_pc[c(rets$up,rets$down),c(Aidx,Bidx)]+1)
  Ret_cor <- cor(Ret_mat)
  
  lower = 0.2
  upper = 1
  pal <- "Reds"
  
  ht1 <- ComplexHeatmap::Heatmap(Ret_cor, col = circlize::colorRamp2(seq(lower, upper, ((upper-lower)/7)),RColorBrewer::brewer.pal(8, pal)),
                  heatmap_legend_param = list(
                  color_bar = "continuous",
                  legend_direction = "horizontal",
                  legend_width = unit(5, "cm"),
                  title_position = "topcenter"),
                  name = "Pearson correlation",
                  column_names_gp = grid::gpar(fontsize = 10),
                  row_names_gp = grid::gpar(fontsize = 10),
                  top_annotation = NULL)
  heatplot3 <- ComplexHeatmap::draw(ht1, heatmap_legend_side = "top")
``` 

```{r eval=FALSE, include=FALSE}
pdf(file = paste0("./figures.tissues/correlation3.edgeR_DEGs.p0.05_FC1.5.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 5.2, width = 5.2)
heatplot3
dev.off()
```

### volcano plot        

##### p.adjust top40 labeled, volcano         

```{r echo=FALSE, fig.height=12, fig.width=15, warning=FALSE}
rets$vol
```

```{r eval=FALSE, include=FALSE}
pdf(file = paste0("./figures.tissues/volcano_edgeR.top40_label.p0.05_FC1.5.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 12, width = 15)
rets$vol
dev.off()
```


##### volcano mod05220523             

```{r message=FALSE, warning=FALSE, include=FALSE}
rets <- finalplot(mat_pc[,c(Aidx,Bidx)],data.frame(ret$DEG_edgeR), paste0(Aname,"_vs_",Bname), 0.05, 1.5, 
                  Sign = FALSE, Sign_dn = 10, Sign_up = 40, padjust = TRUE)
```

```{r echo=FALSE, fig.height=8, fig.width=10, warning=FALSE}
rets$vol
```

```{r eval=FALSE, include=FALSE}
pdf(file = paste0("./figures.tissues/volcano_edgeR.top40_label.p0.05_FC1.5.mod20220523.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 8, width = 10)
rets$vol
dev.off()
```

### heatmap         

```{r echo=FALSE, fig.height=9.6, fig.width=6.8}
genes <- c(rets$up[1:40],rets$down[1:40])
pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8,
         main = "top40 log2(CPM+1)", gaps_row = c(40), gaps_col = length(Aidx))
```

```{r echo=FALSE, fig.height=9.6, fig.width=6.8}
color.test <- colorRampPalette(
    c(
      "#03047F", # deep blue
      "white",
      "#CC2627"  # deep red
    )
  )(100)

pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8, color = color.test,
         scale = 'row',
         main = "top40 zscore", gaps_row = c(40), gaps_col = length(Aidx))

```

```{r eval=FALSE, include=FALSE}
pdf(file = paste0("./figures.tissues/heatmap_top40_log2cpm.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 7.6, width = 5.8)
pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 7.2,
         main = "top40 log2(CPM+1)", gaps_row = c(40), gaps_col = length(Aidx))
dev.off()

pdf(file = paste0("./figures.tissues/heatmap_top40_zscore.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 7.6, width = 5.8)
pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 7.2, color = color.test,
         scale = 'row',
         main = "top40 zscore", gaps_row = c(40), gaps_col = length(Aidx))
dev.off()
```

```{r eval=FALSE, include=FALSE}
write.table(rec_gcol(data.frame(ret$DEG_edgeR)), paste0("edgeR_DEGs.tissues_a7_",
           #paste(colnames(mat_pc)[Aidx],collapse="."),"_vs_", paste(colnames(mat_pc)[Bidx],collapse="."),".csv"),
           paste(Aname),"_vs_", paste(Bname),".csv"),
           col.names = TRUE, row.names = FALSE,quote=FALSE, sep=",")
```


##### mod20220523            

```{r echo=FALSE, fig.height=9.6, fig.width=6.8}
color.test <- colorRampPalette(
    c(
      "#03047F", # deep blue
      "white",
      "#CC2627"  # deep red
    )
  )(100)

pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8, color = color.test,
         scale = 'row', border_color = NA, 
         breaks = seq(-2.5,2.5,0.05),
         main = "top40 zscore", gaps_row = c(40), gaps_col = length(Aidx))

```


```{r echo=FALSE, fig.height=9.6, fig.width=6.8}
color.test <- colorRampPalette(
    c(
      "#03047F", # deep blue
      "white",
      "#CC2627"  # deep red
    )
  )(100)

pheatmap::pheatmap(log2(mat_pc[c(rets$up,rets$down),
                               c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8, color = color.test, show_rownames = F,
         scale = 'row',
         breaks = seq(-2.5,2.5,0.05), border_color = NA,
         main = paste0("zscore of all DEGs(p.adj<0.05, |FC|>1.5)\nSI up: ",length(rets$up),", nonSI up: ",length(rets$down)), gaps_row = c(length(rets$up)), gaps_col = length(Aidx))

```

```{r eval=FALSE, include=FALSE}
pdf(file = paste0("./figures.tissues/heatmap_top40_zscore.mod0523.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 7.6, width = 5.4)
pheatmap::pheatmap(log2(mat_pc[genes,c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 7.2, color = color.test,
         scale = 'row', border_color = NA, 
         breaks = seq(-2,2,0.04),
         main = "top40 zscore", gaps_row = c(40), gaps_col = length(Aidx))
dev.off()

pdf(file = paste0("./figures.tissues/heatmap_DEGs_padj0.05_FC1.5_zscore.mod0523.",
                  Aname,"_vs_",Bname,".pdf"),
    height = 7.6, width = 4.8)
pheatmap::pheatmap(log2(mat_pc[c(rets$up,rets$down),
                               c(Aidx,Bidx)]+1),cluster_rows = F, cluster_cols = F,
         fontsize_row = 8, color = color.test, show_rownames = F,
         scale = 'row',
         breaks = seq(-2,2,0.04), border_color = NA,
         main = paste0("zscore of all DEGs(p.adj<0.05, |FC|>1.5)\nSI up: ",length(rets$up),", nonSI up: ",length(rets$down)), gaps_row = c(length(rets$up)), gaps_col = length(Aidx))
dev.off()
```

























